---
title: "R Notebook"
output: html_notebook
---

The source is from Baidu https://qianxi.baidu.com/. 
1. On the Baidu Qianxi website, it just lists the top 100 cities and values indicate the percentages.
2. For baidu_in_.csv, the columns are the in-cities; for baidu_out.csv, the columns are the out-cities.
3. January 1 ~ March 8, the city names are in Chinese; after March 9, the city naems are in English
4. Please read 'Index_City_EN_CH.csv' file to convert the Chinese and English names
5. After April 5, city index added to the names. The order of index and column names keep the same in all files.  


***If you use the data, please cite the data source and our dataset https://doi.org/10.7910/DVN/FAEZIO.

China Data Lab, 2020, "Baidu Mobility Data", https://doi.org/10.7910/DVN/FAEZIO, Harvard Dataverse, V15, UNF:6:7UstLboGeHwlEYpNmILncg== [fileUNF]

```{r}
library(tidyverse)
```

```{r}
baiduIn <- read.csv("baidu_in_clean.csv")
baiduIn <- mutate(baiduIn, date = as.Date(Date,"%m/%d/%y"))
```

```{r}
baiduIn %>% 
  filter(City_in == "Beijing_110000") %>%
  group_by(City_out)%>%
  summarise(avg = mean(X.,na.rm=T)) %>%
  na.omit() %>%
  arrange(desc(avg))%>%
  slice(11:20) %>%
  mutate(City_out_name = substr(as.character(City_out),1,nchar(as.character(City_out))-7)) %>%
  ggplot()+
  geom_bar(mapping = aes(x = City_out_name, y = avg), stat = "identity",fill="steelblue") +
  theme_minimal()
```

```{r}
baiduOut <- read.csv("baidu_out_clean.csv")
baiduOut <- baiduOut %>% 
  mutate(date = as.Date(Date,"%m/%d/%y")) %>%
  mutate(city_out = as.character(City_out)) %>%
  mutate(city_in = as.character(City_in)) %>%
  select(date,city_out,city_in,X.)
```


```{r}
baiduOut %>% 
  filter(city_out == "Shanghai_310000",city_in == "Wuhan_420100") %>%
  ggplot(mapping = aes(x = date,y = X.))+
  geom_point()
```

```{r}
baiduOut %>% 
  filter(city_out == "Wuhan_420100",city_in == "Shanghai_310000") %>%
  ggplot(mapping = aes(x = date,y = X.))+
  geom_point()
```

```{r}
baiduOut %>% 
  filter(city_out == "Shanghai_310000") %>%
  group_by(city_in)%>%
  summarise(avg = mean(X.,na.rm=T)) %>%
  na.omit() %>%
  arrange(desc(avg))%>%
  slice(1:5) %>%
  mutate(city_in_name = substr(as.character(city_in),1,nchar(as.character(city_in))-7)) %>%
  ggplot()+
  geom_bar(mapping = aes(x = city_in_name, y = avg), stat = "identity",fill="steelblue") +
  theme_minimal()
```

```{r}
chunyun <- read.csv("data/Mobility_chunyun2018.csv")
chunyun
```
```{r}
chunyuncode <- read.csv("chunyunFinal.csv")
chunyuncode <- chunyuncode %>% 
  mutate(date = as.Date(date,"%m/%d/%y")) %>%
  mutate(origin = as.character(Origin)) %>%
  mutate(destination  = as.character(Destination)) %>%
  select(Day,origin,destination,Mobility.Index,date)
chunyuncode
```

```{r}
chunyun %>%
  filter(Origin == "Beijing", Destination == "Shanghai") %>%
  ggplot()+
  geom_point(mapping = aes(x = Day, y = Mobility.Index))
```

```{r}
baiduOut %>%
  filter(city_out == "Beijing_110000", city_in == "Shanghai_310000", date > as.Date("2020-01-09"), date < as.Date("2020-01-24")) %>%
  ggplot()+
  geom_point(mapping = aes(x = date, y = X.))
```

```{r}
chunyun %>%
  group_by(Day,Origin) %>%
  summarise(total_out = sum(Mobility.Index)) %>%
  arrange(Day,desc(total_out)) %>%
  filter(Origin == "Beijing")
```
```{r}
beijingDay1 <- chunyun%>%
  filter(Origin == "Beijing",Day == 1)
beijingDay1
```

```{r}
processCY <- function(cityOut){
  dat <- filter(chunyuncode,origin == cityOut)
  dat$destination
}
```

Wenzhou data is missing here. 
```{r}
baiduOut %>%
  filter(city_out == "Beijing_110000", substr(as.character(city_in),1,nchar(as.character(city_in))-7) %in% beijingDay1$destination, date > as.Date("2020-01-09"), date < as.Date("2020-01-24")) %>%
  na.omit() %>%
  group_by(date) %>%
  summarise(sumPercent = sum(X.))
```

```{r}
totalBaidu <- baiduOut %>%
  filter(date > as.Date("2020-01-09"), date < as.Date("2020-01-24")) %>%
  group_by(city_out) %>%
  filter(as.character(city_in) %in% processCY(as.character(city_out))) %>%
  na.omit() %>%
  group_by(date,city_out) %>%
  summarise(sumPercent = sum(X.))
totalBaidu
```
```{r}
totalBaidu %>% 
  filter(sumPercent > 0) %>%
  group_by(city_out) %>%
  summarise(avgPercent = mean(sumPercent)) %>%
  arrange(avgPercent)
  # ggplot()+
  # geom_point(aes(x = city_out,y = avgPercent))
```

```{r}
sumCY <- chunyuncode %>%
  group_by(date,origin) %>%
  summarise(total_out = sum(Mobility.Index)) %>%
  arrange(date) 
sumCY
```
It seems that the data for Hainan is very incomplete during that period in 2020. 

```{r}
combo <- inner_join(sumCY, totalBaidu, by = c("origin" = "city_out","date" = "date"))
combo <- filter(combo,total_out != 0, sumPercent != 0)
```

```{r}
combo <- combo %>% 
  mutate(total_est = total_out/(sumPercent/100))
  # select(date,Origin,total_est) 
  # filter(Origin == "Shanghai_310000") %>%
  # group_by(date) %>%
  # summarise(sum_out = sum(total_est))
combo
```

```{r}
rand <- sample_n(combo,10) %>% filter(date == as.Date("2020-01-10"))
rand$origin
```

```{r}
combo %>% 
  filter(origin %in% rand$origin) %>%
  ggplot()+
  geom_line(mapping = aes(x = date, y = total_est,color = origin))
```

```{r}
sumCY %>%
  filter(origin %in% rand$origin) %>%
  ggplot()+
  geom_line(mapping = aes(x = date, y = total_out,color = origin))
```

```{r}
exp <- c("Beijing_110000","Shanghai_310000","Wuhan_420100")
combo %>% 
  filter(origin %in% exp) %>%
  ggplot()+
  geom_line(mapping = aes(x = date, y = total_est,color = origin))
```

```{r}
sumCY%>% 
  filter(origin %in% exp) %>%
  ggplot()+
  geom_line(mapping = aes(x = date, y = total_out,color = origin))
```

```{r}
chunyuncode %>%
  filter(origin == "Beijing_110000", destination == "Shanghai_310000") %>%
  ggplot()+
  geom_line(mapping = aes(x = Day, y = Mobility.Index))
```

```{r}
write.csv(combo,"predicted_total_outflow.csv")
```

```{r}
baiduCY <- baiduOut %>%
  filter(date > as.Date("2020-01-09"), date < as.Date("2020-01-24")) %>%
  na.omit() %>%
  filter(X. != 0)
baiduCY
```

```{r}
write.csv(baiduCY,"baidu_chunyun.csv")
```

```{r}
newBaiduout <- read.csv("new_baidu_out.csv")
newBaiduout <- newBaiduout %>%
  select(-Unnamed..0) %>%
  mutate(date = as.Date(date))
newBaiduout
```

```{r}
newBaiduout %>%
  filter(city_out == "Beijing_110000", city_in == "Shanghai_310000") %>%
  ggplot()+
  geom_line(mapping = aes(x = date, y = predicted_flow))
```


